Kindly Bent to Free Us

### 2020-05-24 kindly/intro.tex
Worked on 'Typos and Focused Remarks' from Reviewer C.

The design of Rust is geared towards programmers with a low-level
imperative programming background, like C or C++. Its management of
lifetimes supports the manual way of memory management customary in
these languages very well and makes it safe. However, programmers with
a background in managed languages feel alienated from the lack of garbage
collected data. They would prefer a setting where automatic memory
management with garbage collection was the default, but where they
could seemlessly switch to safe, manual resource management if that
was required.
%
As a concrete example, consider a functional programmer who wants to
safely interact with a C library. Invoking a C function is easy via
the existing foreign function interface, but managing the underlying
resources like malloc'd storage is not: It cannot be left to the
garbage collector, but proper release of the storage via calls to \texttt{free()} must be
ensured by programming conventions.

>>>>

The design of Rust is geared towards programmers with a low-level
imperative programming background, like C or C++. Its management of
lifetimes supports the manual way of memory management customary in
these languages very well and makes it safe. However, programmers with
a background in managed languages feel alienated from the lack of garbage
collected data. They would prefer a setting where automatic memory
management with garbage collection was the default, but where they
could seemlessly switch to safe, manual resource management if that
was required.
%
As a concrete example, consider a functional programmer who wants to
safely interact with a C library. Invoking a C function is easy via
the existing foreign function interface, but managing the underlying
resources like malloc'd storage is not: it cannot be left to the
garbage collector, but proper release of the storage via calls to \texttt{free()} must be
ensured by programming conventions.

### 2020-05-24 kindly/lang.tex
instance
Worked on 'Typos and Focused Remarks' from Reviewer C.

In the surface language (\cref{motivation}) region annotations are optional.
In the internal language, regions must be syntactically explicit and
annotated with a nesting index and a scoped variable.
This section defines a transformation $\RannotT{p}{p'}$ which
automatically inserts region annotations in  programs.
The input $p$ is a program with optional region annotations of the form $\region[]{}{e}$.
The output $p'$ is a program with explicit annotations
of the form $\region{\Sone x\BORROW}{e}$ such that no borrow occurs
outside a region.
We give an informal presentation of our code transformation and defer
the complete definition to \cref{appendix:regionannot}. This code transformation
aims to find, for each borrow $\borrow{x}$, the biggest region satisfying
the following rules:
\begin{enumerate}
\item The region should contain at least $\borrow{x}$.
\item The region must be contained in the scope of $x$.
\item An exclusive borrow $\borrow[\kaff]{x}$ should never share a region with any other borrow of $x$.
\item The variable $x$ cannot occur in the region of $\borrow{x}$.
\end{enumerate}
The transformation starts from each borrow $\borrow x$ and grows its associated
region until enlarging it would include the binding for $x$ or lead to a
conflicting use of $x$ or a conflicting borrow for $x$.
As an example, consider the following program:
%
\[
  \setlength{\jot}{-1pt}
  \RannotT{
\begin{aligned}
  \lam{a}{}\ &
  \letin{x}{\app{f}{\borrow[\kun]{a}}}{}\\
  &g\ (\borrow[\kaff]{x});\\
  &f\ (\borrow[\kun]{x})\ (\borrow[\kun]{x})\\
\end{aligned}
}{
\begin{aligned}
  \lam{a}{}\region[1]{\Sone a\MBORROW}{\ &%
  \letin{x}{\app{f}{\borrow[\kaff]{a}}}{}\\
  &\region[2]{\Sone x\MBORROW}{g\ (\borrow[\kaff]{x})};\\
  &\region[2]{\Sone x\IBORROW}{f\ (\borrow[\kun]{x})\ (\borrow[\kun]{x})}\\
  &\!\!\!}
\end{aligned}
}
\]
%
As variable $a$ only has one borrow, its region covers its whole lexical scope.
Variable $x$ has multiple conflicting borrows and requires more consideration.
We place a first
region around the exclusive borrow and its function application,
and a second region around both shared borrows. This placement of region
is optimal: making any region bigger would cause a borrowing
error.
Region indices are assigned after placement, so it is trivial to
ensure well-nested regions where the inner regions have
higher indices than the outer ones.

@@

\begin{figure*}[tp]
  % \begin{minipage}{0.7\linewidth}
    \begin{mathpar}
      \ruleSDVar
      \and
      \ruleSDApp
      \and
      \ruleSDRegion
      \and
      \ruleSDBorrow
      \and
      \ruleSDReBorrow
      \and
      \ruleSDLam
      \and
      \inferrule[BorrowBinding]{
        \entail{C}{(\BORROW_n\lk k) \wedge (k \lk \BORROW_\infty)}\\
        \BORROW\in\left\{\kun,\kaff\right\}
      }{
        \lregion{C}{}
        {\svar[\BORROW]{x}{\schm}^n}
        {\bbvar x k \schm}
        % {\bvar{\borrow[\BORROW]{x}}{\borrowty[\BORROW] k{\schm}}}
      }
      % \and
      % \ruleSDPair
      % \and
      % \ruleSDMatchPair
    \end{mathpar}
    \vspace{-10pt}
    \caption{Selected typing rules ($\inferS{C}{\E}{e}{\tau}$)
      and borrowing rules ($\lregion{C}{x}{\E}{\E'}$)}
    \label{selectrules:borrow}
    \label{selectrules:binders}
    \label{sdtyp:app}
    \label{selectrules:region}
    \label{env:rule:borrow}
    \vspace{-5pt}
  % \end{minipage}
\end{figure*}

@@

Addresses $\Addr$ are composed of a raw location $\Loc$, which is just
a pointer into a store, and a stack of modifiers that indicates the
borrows that have been taken from the original object. From the raw
location, we may take affine borrows and reborrows. Once we have
taken an unrestricted borrow (from a raw location or a borrowed one),
then we can take further unrestricted borrows from it, but no more
affine ones.

>>>>

In the surface language (\cref{motivation}) region annotations are optional.
In the internal language, regions must be syntactically explicit and
annotated with a nesting index and a scoped variable.
This section defines a transformation $\RannotT{p}{p'}$ which
automatically inserts region annotations in  programs.
The input $p$ is a program with optional region annotations of the form $\region[]{}{e}$.
The output $p'$ is a program with explicit annotations
of the form $\region{\Sone x\BORROW}{e}$ such that no borrow occurs
outside a region.
We give an informal presentation of our code transformation and defer
the complete definition to \cref{appendix:regionannot}. This code transformation
aims to find, for each borrow $\borrow{x}$, the biggest region satisfying
the following rules:
\begin{enumerate}
\item The region should contain at least $\borrow{x}$.
\item The region must be contained in the scope of $x$.
\item An exclusive borrow $\borrow[\kaff]{x}$ should never share a region with any other borrow of $x$.
\item The variable $x$ cannot occur in the region of $\borrow{x}$.
\end{enumerate}
The transformation starts from each borrow $\borrow x$ and grows its associated
region until enlarging it would include the binding for $x$ or lead to a
conflicting use of $x$ or a conflicting borrow for $x$.
As an example, consider the following program:
%
\[
  \setlength{\jot}{-1pt}
  \RannotT{
\begin{aligned}
  \lam{a}{}\ &
  \letin{x}{\app{f}{\borrow[\kun]{a}}}{}\\
  &g\ (\borrow[\kaff]{x});\\
  &f\ (\borrow[\kun]{x})\ (\borrow[\kun]{x})\\
\end{aligned}
}{
\begin{aligned}
  \lam{a}{}\region[1]{\Sone a\MBORROW}{\ &%
  \letin{x}{\app{f}{\borrow[\kun]{a}}}{}\\
  &\region[2]{\Sone x\MBORROW}{g\ (\borrow[\kaff]{x})};\\
  &\region[2]{\Sone x\IBORROW}{f\ (\borrow[\kun]{x})\ (\borrow[\kun]{x})}\\
  &\!\!\!}
\end{aligned}
}
\]
%
As variable $a$ only has one borrow, its region covers its whole lexical scope.
Variable $x$ has multiple conflicting borrows and requires more consideration.
We place a first
region around the exclusive borrow and its function application,
and a second region around both shared borrows. This placement of region
is optimal: making any region bigger would cause a borrowing
error.
Region indices are assigned after placement, so it is trivial to
ensure well-nested regions where the inner regions have
higher indices than the outer ones.

@@

\begin{figure*}[tp]
  % \begin{minipage}{0.7\linewidth}
    \begin{mathpar}
      \ruleInstance
      \and
      \ruleSDVar
      \and
      \ruleSDLam
      \and
      \ruleSDApp
      \and
      \ruleSDRegion
      \and
      \ruleSDBorrow
      \and
      \ruleSDReBorrow
      \and
      \inferrule[BorrowBinding]{
        \entail{C}{(\BORROW_n\lk k) \wedge (k \lk \BORROW_\infty)}\\
        \BORROW\in\left\{\kun,\kaff\right\}
      }{
        \lregion{C}{}
        {\svar[\BORROW]{x}{\schm}^n}
        {\bbvar x k \schm}
        % {\bvar{\borrow[\BORROW]{x}}{\borrowty[\BORROW] k{\schm}}}
      }
      % \and
      % \ruleSDPair
      % \and
      % \ruleSDMatchPair
    \end{mathpar}
    \vspace{-10pt}
    \caption{Selected typing rules ($\inferS{C}{\E}{e}{\tau}$)
      and borrowing rules ($\lregion{C}{x}{\E}{\E'}$)}
    \label{selectrules:borrow}
    \label{selectrules:binders}
    \label{sdtyp:app}
    \label{selectrules:region}
    \label{env:rule:borrow}
    \vspace{-5pt}
  % \end{minipage}
\end{figure*}

@@

Addresses $\Addr$ are composed of a raw location $\Loc$, which is just
a pointer into a store, and a stack of modifiers that indicates the
borrows and reborrows that have been taken from $\Loc$.
Once we have taken an unrestricted borrow (from a raw location or a borrowed
one), then we can take further unrestricted borrows from it, but no more affine
ones.
